### 2017-01-22 previous-paper-analysis/previous-paper-analysis.tex
Starting previous paper analysis

\documentclass[12pt,letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}

\title{Previous Paper Analysis}
\date{January 24, 2017}
\author{Shawn Seymour}

\begin{document}

\maketitle

\newpage

\begin{thebibliography}{9}

\bibitem{grove}
	Molly Grove,
	\emph{Heuristics for the Generalized Traveling Salesman Problem},
	UMM CSci Senior Seminar Conference, Fall 2015, Morris, MN.

\end{thebibliography}

\end{document}