### 2017-04-15 presentation/presentation.tex
Last minute changes

\documentclass[xcolor=dvipsnames,aspectratio=1610]{beamer}

\usepackage{graphicx}
\usepackage[scale=2]{ccicons}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{algorithm, caption}
\usepackage[noend]{algpseudocode}
\usepackage[absolute,overlay]{textpos}
\usepackage{calc}
\usepackage[backend=bibtex,citestyle=authortitle]{biblatex}
\usepackage[no-math]{fontspec}

% \usepackage[texcoord,grid,gridunit=mm,gridcolor=red!10,subgridcolor=green!10]
% {eso-pic}

\setitemize{label=\usebeamerfont*{itemize item}%
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}

\newcommand{\exampleheight}{1.9cm}
\newcommand{\examplewidth}{16cm}

% \makeatletter
% \newcommand{\algcolor}[2]{%
%   \hskip-\ALG@thistlm\colorbox{#1}{\raisebox{0pt}[0.1cm][0cm]{\parbox{\dimexpr\linewidth-2\fboxsep}{\hskip\ALG@thistlm\relax #2}}}%
% }
% \newcommand{\algemph}[1]{\algcolor{Yellow}{#1}}
% \makeatother

\usepackage{tikz}
\usetikzlibrary{calc}

\makeatletter
% code borrowed from Andrew Stacey; See
% http://tex.stackexchange.com/a/50054/3954
\tikzset{%
  remember picture with id/.style={%
    remember picture,
    overlay,
    save picture id=#1,
  },
  save picture id/.code={%
    \edef\pgf@temp{#1}%
    \immediate\write\pgfutil@auxout{%
      \noexpand\savepointas{\pgf@temp}{\pgfpictureid}}%
  },
  if picture id/.code args={#1#2#3}{%
    \@ifundefined{save@pt@#1}{%
      \pgfkeysalso{#3}%
    }{
      \pgfkeysalso{#2}%
    }
  }
}

\def\savepointas#1#2{%
  \expandafter\gdef\csname save@pt@#1\endcsname{#2}%
}

\def\tmk@labeldef#1,#2\@nil{%
  \def\tmk@label{#1}%
  \def\tmk@def{#2}%
}

\tikzdeclarecoordinatesystem{pic}{%
  \pgfutil@in@,{#1}%
  \ifpgfutil@in@%
    \tmk@labeldef#1\@nil
  \else
    \tmk@labeldef#1,(0pt,0pt)\@nil
  \fi
  \@ifundefined{save@pt@\tmk@label}{%
    \tikz@scan@one@point\pgfutil@firstofone\tmk@def
  }{%
  \pgfsys@getposition{\csname save@pt@\tmk@label\endcsname}\save@orig@pic%
  \pgfsys@getposition{\pgfpictureid}\save@this@pic%
  \pgf@process{\pgfpointorigin\save@this@pic}%
  \pgf@xa=\pgf@x
  \pgf@ya=\pgf@y
  \pgf@process{\pgfpointorigin\save@orig@pic}%
  \advance\pgf@x by -\pgf@xa
  \advance\pgf@y by -\pgf@ya
  }%
}

\makeatother
% end of Andrew's code

% main command to draw the colored background
\newcounter{mymark}
\newcommand\ColorLine{%
  \stepcounter{mymark}%
  \tikz[remember picture with id=mark-\themymark,overlay] {;}%
  \begin{tikzpicture}[remember picture,overlay]%
    \filldraw[Yellow]%
   let \p1=(pic cs:mark-\themymark),
   \p2=(current page.east)  in
   ([xshift=-0.3em,yshift=-0.7ex]0,\y1)  rectangle ++([xshift=-8.26cm]\x2,\baselineskip);
  \end{tikzpicture}%
}%

% colored loops and declarations
\makeatletter
\algnewcommand\CREQUIRE{\item[\ColorLine\algorithmicrequire]}%
\algnewcommand\CENSURE{\item[\ColorLine\algorithmicensure]}%
\algnewcommand\CSTATE{\State\ColorLine}%
\algnewcommand\CSTATEx{\Statex\ColorLine}%
\algnewcommand\CCOMMENT{\Comment\ColorLine}%

% \algdef{SE}[WHILE]{CWhile}{EndWhile}%
%    [2][default]{\ColorLine\algorithmicwhile\ #2\ \algorithmicdo\ALG@compatcomm{#1}}%
%    {\algorithmicend\ \algorithmicwhile}%
 \algdef{SE}[WHILE]{CWhile}{EndWhile}[1]{\ColorLine\algorithmicwhile\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicwhile}%

% \algdef{SE}[FOR]{CFOR}{ENDFOR}%
%    [2][default]{\ColorLine\algorithmicfor\ #2\ \algorithmicdo\ALG@compatcomm{#1}}%
%    {\algorithmicend\ \algorithmicfor}%
% \algdef{S}[FOR]{CFORALL}%
%    [2][default]{\ColorLine\algorithmicforall\ #2\ \algorithmicdo\ALG@compatcomm{#1}}%

\algdef{SE}[FOR]{For}{EndFor}[1]{\algorithmicfor\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicfor}%

\algdef{S}[FOR]{CForAll}[1]{\ColorLine\algorithmicforall\ #1\ \algorithmicdo}%

\algdef{SE}[LOOP]{CLOOP}{ENDLOOP}%
   [1][default]{\ColorLine\algorithmicloop\ALG@compatcomm{#1}}%
   {\algorithmicend\ \algorithmicloop}%
\algdef{SE}[REPEAT]{CREPEAT}{UNTIL}%
   [1][default]{\ColorLine\algorithmicrepeat\ALG@compatcomm{#1}}%
   [1]{\algorithmicuntil\ #1}%
\algdef{SE}[IF]{CIF}{ENDIF}%
   [2][default]{\ColorLine\algorithmicif\ #2\ \algorithmicthen\ALG@compatcomm{#1}}%
   {\algorithmicend\ \algorithmicif}%
\algdef{C}[IF]{IF}{CELSIF}%
   [2][default]{\ColorLine\algorithmicelse\ \algorithmicif\ #2\ \algorithmicthen\ALG@compatcomm{#1}}%
\algdef{Ce}[ELSE]{IF}{CELSE}{ENDIF}%
   [1][default]{\ColorLine\algorithmicelse\ALG@compatcomm{#1}}%
\makeatother
\algtext*{EndWhile}%
\algtext*{EndFor}%

\usetheme[numbering=counter, progressbar=frametitle, sectionpage=none]{metropolis}
\setbeamercolor{background canvas}{bg=white}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\setsansfont[ItalicFont={Fira Sans Light Italic},%
             BoldFont={Fira Sans Bold},%
             BoldItalicFont={Fira Sans Italic}]%
            {Fira Sans Light}

\title{Conflict-Free Vertex Coloring of Planar Graphs}
\date{April 15, 2017}
\author{Shawn Seymour}

\begin{document}
  \maketitle

  \begin{frame}
    \begin{figure}[h]
      \centering
      \includegraphics[width=14cm]{../figures/map-no-colors-1.pdf}
    \end{figure}
  \end{frame}

  \begin{frame}
    \begin{figure}[h]
      \centering
      \includegraphics[width=14cm]{../figures/map-colors.pdf}
    \end{figure}
  \end{frame}

  \begin{frame}
    \frametitle{Overview}
    \begin{multicols}{2}
      \tableofcontents
    \end{multicols}
  \end{frame}

  \addtocontents{toc}{\protect\vspace{-14pt}}

  \section{Background}

  \subsection{Graph Theory}

  \begin{frame}
    \frametitle{Graph Theory}

    \only<1-3,6,9,11>{
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example.pdf}
      \end{textblock*}
    }

    \only<4> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-neighborhoods-1.pdf}
      \end{textblock*}
    }

    \only<5> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-neighborhoods-2.pdf}
      \end{textblock*}
    }

    \only<7> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-dist-1.pdf}
      \end{textblock*}
    }

    \only<8> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-dist-2.pdf}
      \end{textblock*}
    }

    \only<10> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-d3s.pdf}
      \end{textblock*}
    }

    \only<12> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-dom-1.pdf}
      \end{textblock*}
    }

    \only<13> {
      \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=8cm]{../figures/example-dom-2.pdf}
      \end{textblock*}
    }

    \vspace{3.5cm}

    \vfill

    \only<1-2>{
      \begin{itemize}
        \item \textbf{Simple graph}: undirected graph with no loops.
        \pause
        \vfill
        \item \textbf{Planar graph}: no edges cross.
      \end{itemize}
    }

    \only<3-5>{
      \begin{itemize}
        \item<3-5> \textbf{Neighborhood} of a vertex: a set of all adjacent vertices and the vertex itself.
        \vfill
        \only<3>{\item<4> Example: The neighborhood of vertex 1: $\{1, 2, 5\}$.}
        \only<4>{\item<4> Example: The neighborhood of vertex 1: $\{1, 2, 5\}$.}
        \only<5>{\item<5> Example: The neighborhood of vertex 5: $\{1, 2, 4, 5, 6\}$.}
      \end{itemize}
    }

    \only<6-8>{
      \begin{itemize}
        \item \textbf{Distance}: smallest number of edges to get from one vertex to another.
        \vfill
        \only<6>{\item<7> Example: distance from vertex 1 to 3: \textbf{2}.}
        \only<7>{\item<7> Example: distance from vertex 1 to 6: \textbf{2}.}
        \only<8>{\item<8> Example: distance from vertex 3 to 4: \textbf{3}.}
      \end{itemize}
    }

    \only<9-10>{
      \begin{itemize}
        \item \textbf{Distance-3-set}: contains all vertices with exactly distance 3 from each other.
        \vfill
        \only<9>{\item<7> Example: Distance from vertex 1 to 3: \textbf{2}.}
        \only<10>{\item<10> Example: The only possible distance-3-set: $\{3, 4\}$.}
      \end{itemize}
    }

    \only<11-13>{
    \vspace{0.55cm}
      \begin{itemize}
        \item \textbf{Dominating set}: all vertices not in the set must have distance 1 to some vertex within the set.
        \vfill
        \only<11>{\item<7> Example: The neighborhood of vertex 1: $\{1, 2, 5\}$.}
        \only<12>{\item<12> Example: $\{2, 5\}$}
        \only<13>{\item<13> Example: $\{2, 4\}$}
      \end{itemize}
    }

  \end{frame}

  \subsection{Vertex Coloring}

  \begin{frame}
    \frametitle{Vertex Coloring}

    \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
      \centering
      \includegraphics[width=8cm]{../figures/example-vcp.pdf}
    \end{textblock*}

    \vspace{4.5cm}
    \vfill

    \begin{itemize}
      \item A \textbf{proper vertex coloring} assigns colors to \emph{every} vertex such that no two adjacent vertices share the same color.
      \pause
      \vfill
      \item The \textbf{chromatic number} is the minimum number of colors needed to properly color a graph.
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Conflict-Free Coloring}

    \begin{textblock*}{\examplewidth}(0cm,\exampleheight) % {block width} (coords)
      \centering
      \includegraphics[width=8cm]{../figures/example-cfcp.pdf}
    \end{textblock*}

    \vspace{4.5cm}
    \vfill

    \begin{itemize}
      \item<1-2> A \textbf{conflict-free coloring} assigns colors to \emph{some} vertices such that the neighborhood of every vertex contains at least one uniquely colored vertex.
      \vfill
      \pause
      \item<2> Proper vertex colorings are also conflict-free colorings.
    \end{itemize}
  \end{frame}

  \begin{frame}
    \frametitle{Conflict-Free Coloring}

    \only<1, 8->{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cfcp.pdf}
      \end{textblock*}
    }

    \only<2>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-1.pdf}
      \end{textblock*}
    }

    \only<3>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-2.pdf}
      \end{textblock*}
    }

    \only<4>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-3.pdf}
      \end{textblock*}
    }

    \only<5>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-4.pdf}
      \end{textblock*}
    }

    \only<6>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-5.pdf}
      \end{textblock*}
    }

    \only<7>{
      \begin{textblock*}{\examplewidth/2}(0cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-cf-6.pdf}
      \end{textblock*}
    }

    \only<1-7>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp.pdf}
      \end{textblock*}
    }

    \only<8>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-1.pdf}
      \end{textblock*}
    }

    \only<9>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-2.pdf}
      \end{textblock*}
    }

    \only<10>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-3.pdf}
      \end{textblock*}
    }

    \only<11>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-4.pdf}
      \end{textblock*}
    }

    \only<12>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-5.pdf}
      \end{textblock*}
    }

    \only<13>{
      \begin{textblock*}{\examplewidth/2}(8cm,\exampleheight) % {block width} (coords)
        \centering
        \includegraphics[width=6.5cm]{../figures/example-vcp-6.pdf}
      \end{textblock*}
    }

    \vspace{3cm}

    \pause

    \begin{columns}
      \begin{column}{0.5\textwidth}
        \begin{multicols}{2}
          \begin{itemize}[leftmargin=1.4cm]
            \item[1:] \textbf{red}
            \pause
            \item[2:] \textbf{blue}
            \pause
            \item[3:] \textbf{blue}
            \pause
            \item[4:] \textbf{red}
            \pause
            \item[5:] \textbf{red}
            \pause
            \item[6:] \textbf{red}
          \end{itemize}
        \end{multicols}
      \end{column}

      \pause

      \begin{column}{0.5\textwidth}
        \begin{multicols}{2}
          \begin{itemize}[leftmargin=1.4cm]
            \item[1:] \textbf{blue}
            \pause
            \item[2:] \textbf{red}
            \pause
            \item[3:] \textbf{blue}
            \pause
            \item[4:] \textbf{red}
            \pause
            \item[5:] \textbf{yellow}
            \pause
            \item[6:] \textbf{blue}
          \end{itemize}
        \end{multicols}
      \end{column}
    \end{columns}

  \end{frame}

  \subsection{Conflict-Free Coloring}

  \begin{frame}
    \frametitle{Conflict-Free Coloring Examples}

    \begin{textblock*}{\examplewidth}(1cm,\exampleheight) % {block width} (coords)
      Incorrect conflict-free colorings and their fixes:
    \end{textblock*}

    \vspace{0.5cm}

    \pause

    \begin{columns}
      \begin{column}{0.33333\textwidth}
        \centering
        \vspace{4.5cm}
        Isolated vertex
        \only<2-7>{
          \begin{textblock*}{\examplewidth/3}(0.15cm,3.5cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-1.pdf}
          \end{textblock*}
        }

        \only<3-7>{
          \begin{textblock*}{\examplewidth/3}(0.15cm,6.7cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-1-corrected.pdf}
          \end{textblock*}
        }

      \end{column}

      \pause
      \pause

      \begin{column}{0.33333\textwidth}
        \centering
        \vspace{4.5cm}
        Triangle

        \only<4-7>{
          \begin{textblock*}{\examplewidth/3}(\examplewidth/3,3.5cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-2.pdf}
          \end{textblock*}
        }

        \only<5-7>{
          \begin{textblock*}{\examplewidth/3}(\examplewidth/3,6.7cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-2-corrected.pdf}
          \end{textblock*}
        }

      \end{column}

      \pause
      \pause

      \begin{column}{0.33333\textwidth}
        \centering
        \vspace{4.5cm}
        Missing unique

        \only<6-7>{
          \begin{textblock*}{\examplewidth/3}(10.57cm,3.5cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-3.pdf}
          \end{textblock*}
        }

        \only<7>{
          \begin{textblock*}{\examplewidth/3}(10.57cm,6.7cm) % {block width} (coords)
            \centering
            \includegraphics[width=3cm]{../figures/examples-1-incorrect-3-corrected.pdf}
          \end{textblock*}
        }
      \end{column}
    \end{columns}

  \end{frame}

  % \addtocontents{toc}{\protect\vspace{14pt}}

  \subsection{Applications}

  \addtocontents{toc}{\protect\vspace{21pt}}

  \begin{frame}
    \frametitle{Applications}

    \begin{itemize}
      \item \textbf{Applications}: wireless networks, satellite communication systems, RFID networks.

      \pause
      \vfill

      \begin{figure}[h]
        \centering
        \includegraphics[width=12cm]{../figures/towers-slides.pdf}
      \end{figure}

      \vfill

      \item \textbf{Cellular networks}: consist of towers and clients.
      \pause
      \item The problem: \textbf{frequency assignment}.

    \end{itemize}

  \end{frame}

  \begin{frame}
    \frametitle{Cellular Networks}

    \only<1>{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/towers-slides-2.pdf}
      \end{figure}
    }

    \only<2->{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/towers-slides-3.pdf}
      \end{figure}
    }

    \pause
    \pause

    \begin{itemize}
      \item Our goal is to assign frequencies such that:
      \pause
      \begin{itemize}
        \item[(1)] Every client is served by a tower with a unique frequency.
        \pause
        \item[(2)] Minimize the number of frequencies used.
      \end{itemize}
    \end{itemize}

  \end{frame}

  \begin{frame}
    \frametitle{Cellular Networks}

    \begin{figure}[h]
      \centering
      \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/towers-graph.pdf}
    \end{figure}

    \begin{figure}[h]
      \centering
      \includegraphics[width=10cm,trim=4 4 4 4,clip]{../figures/towers.pdf}
      \caption*{Vertex coloring and conflict-free coloring of cellular towers, respectively}
    \end{figure}

  \end{frame}

  % \subsection{RFID Networks}
  %
  % \begin{frame}
  %   \frametitle{RFID Networks}
  % \end{frame}

  %

  \addtocontents{toc}{\newpage}

  \section{Coloring for General Graphs}
  \subsection{Guaranteeing k-Colorability}
  \begin{frame}
    \frametitle{Guaranteeing Conflict-Free k-Colorability}

    We want to guarantee a graph can be colored with $k$ colors.

    \pause

    \only<1-2>{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/bad-criterion-1.pdf}
        \caption*{Complete 3 vertex and complete 4 vertex graphs}
      \end{figure}
    }

    \only<3>{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/bad-criterion-2.pdf}
        \caption*{Complete 3 vertex and complete 4 vertex graphs}
      \end{figure}
    }

    \only<4>{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/bad-criterion.pdf}
        \caption*{Complete 3 vertex and tattered 4 vertex graphs}
      \end{figure}
    }

    \vfill

    \begin{itemize}
      \item \textbf{Complete graph}: every pair of distinct vertices is connected by an edge.
      \vfill
      \pause
      \item \textbf{Tattered graph}: complete graph with a triangle removed.
    \end{itemize}

  \end{frame}

  % \begin{theorem}
  % Let $G$ be a graph and $k \geq 1$. If $G$ has neither $K_{k+2}$ nor $K_{k+3}^{-3}$ as a minor, $G$ has a conflict-free coloring that can be found in polynomial time.
  % \end{theorem}

  \begin{frame}
    \frametitle{Guaranteeing Conflict-Free k-Colorability}

    % \begin{textblock*}{\examplewidth}(1cm,1.4cm) % {block width} (coords)
    % \end{textblock*}

    \begin{textblock*}{14cm}(1cm, 1.6cm) % {block width} (coords)
      Let's demonstrate meeting this criterion on a graph for the simplest case, 1 color.
    \end{textblock*}

    \begin{textblock*}{15cm}(0.5cm, 2.1cm) % {block width} (coords)
      \only<1>{
        \begin{figure}
         \centering
         \includegraphics[width=6cm]{../figures/criterion.pdf}
         \caption*{A simple graph}
       \end{figure}
      }

      \only<2>{
        \vspace{0.25cm}
        \begin{columns}
          \begin{column}{0.5\textwidth}
            \begin{center}
              \begin{figure}
                 \centering
                 \includegraphics[width=6cm]{../figures/criterion.pdf}
                 \caption*{A simple graph}
               \end{figure}
            \end{center}
          \end{column}
          \begin{column}{0.5\textwidth}
            \begin{center}
              \begin{figure}
                \centering
                \includegraphics[width=6cm]{../figures/bad-criterion.pdf}
                \caption*{Complete 3 vertex and tattered 4 vertex graphs}
              \end{figure}
            \end{center}
          \end{column}
        \end{columns}
      }

      \only<3>{
        \vspace{0.25cm}
        \begin{columns}
          \begin{column}{0.5\textwidth}
            \begin{center}
              \begin{figure}
                 \centering
                 \includegraphics[width=6cm]{../figures/criterion-minor.pdf}
                 \caption*{A simple graph}
               \end{figure}
            \end{center}
          \end{column}
          \begin{column}{0.5\textwidth}
            \begin{center}
              \begin{figure}
                \centering
                \includegraphics[width=6cm]{../figures/bad-criterion.pdf}
                \caption*{Complete 3 vertex and tattered 4 vertex graphs}
              \end{figure}
            \end{center}
          \end{column}
        \end{columns}
      }

      \only<4>{
        \begin{textblock*}{14cm}(1cm, 3.5cm) % {block width} (coords)
          \begin{theorem}
          A graph cannot be conflict-free colored with $k$ colors if it contains a complete graph of $k+2$ vertices or a tattered graph of $k+3$ vertices as a minor.
          \end{theorem}
        \end{textblock*}
      }
    \end{textblock*}

    \begin{textblock*}{14cm}(1cm, 7.1cm) % {block width} (coords)
      A graph is a \textbf{minor} if it can be formed by deleting and/or contracting edges from its parent graph.  \only<3->{$G$ contains both graphs as a minor and thus {\color{red}\textbf{cannot}} be conflict-free colored with 1 color.}
    \end{textblock*}

  \end{frame}

  \subsection{Distance-3-Sets Algorithm}

  \begin{frame}
    \frametitle{Iterated Elimination of Distance-3-Sets}

    \input{ieds}

  \end{frame}

  \addtocontents{toc}{\protect\vspace{7pt}}

  \section{Coloring for Planar Graphs}

  \subsection{Bounds for Planar Graphs}
  \begin{frame}
    \frametitle{Bounds on Planar Graphs}

    \only<3>{
      \begin{figure}[h]
        \centering
        \includegraphics[width=8cm,trim=4 4 4 4,clip]{../figures/four.pdf}
        \caption*{A vertex and conflict-free coloring, respectively}
      \end{figure}
    }

    \begin{theorem}
    Every loopless planar graph admits a proper \textbf{vertex coloring} with at most {\color{red} four} distinct colors.
    \end{theorem}

    \pause

    \begin{theorem}
    Every loopless planar graph admits a \textbf{conflict-free coloring} with at most {\color{red} three} distinct colors.
    \end{theorem}

  \end{frame}

  \subsection{Dominating Set Algorithm}
  \begin{frame}
    \frametitle{Conflict-Free Coloring via Dominating Set}

    \input{dominating}
  \end{frame}

  \begin{frame}
    \frametitle{Future Work}

    \begin{itemize}
      \item Finding bounds and properties on specific graphs such as outerplanar graphs, interval graphs, hypergraphs, and more.
      \pause
      \vfill
      \begin{itemize}
        \item Allows for accurate estimates when applying conflict-free coloring to real-world problems.
      \end{itemize}
      \pause
      \vfill
      \item Variations of conflict-free coloring such as requiring \textbf{another} vertex to have a unique color within the neighborhood of a selected vertex.
      \pause
      \vfill
      \begin{itemize}
        \item Guiding a robot (unique color 1) to a destination (unique color 2).
      \end{itemize}
    \end{itemize}

  \end{frame}

  \begin{frame}[standout]
    \centering
    {Thanks to Peter Dolan, Elena Machkasova,

    and Peh Ng for their advice and feedback.}
    \vfill
    \href{https://github.com/devshawn/senior-seminar}{github.com/devshawn/senior-seminar}
    \vfill
    \ccbyncsa{}
  \end{frame}

\end{document}